---
title: "mouse_cellranger"
output: html_document
date: "2023-07-25"
---

```{r setup, include=FALSE}
library(SingleCellExperiment)
library(DropletUtils)
library(ggplot2)
library(scuttle)
library(scater)
library(scran)
library(bluster) 
library(Seurat)
```

## Check variance on cellranger samples
```{r}
sample1 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409869_78/filtered_feature_bc_matrix",'sample1')
sample2 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409870_80/filtered_feature_bc_matrix",'sample2')
sample3 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409871_81/filtered_feature_bc_matrix",'sample3')
sample4 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409872_82/filtered_feature_bc_matrix",'sample4')
sample5 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409873_83/filtered_feature_bc_matrix",'sample5')
sample6 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409874_84/filtered_feature_bc_matrix",'sample6')
sample7 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409875_85/filtered_feature_bc_matrix",'sample7')
sample8 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409876_86/filtered_feature_bc_matrix",'sample8')
sample9 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409877_87/filtered_feature_bc_matrix",'sample9')
```

```{r}
# Assume your count matrices are called "counts1", "counts2", ..., "counts9"

# Combine matrices into a list
counts_list <- list(sample1$nFeature_RNA, sample2$nFeature_RNA, sample3$nFeature_RNA, sample4$nFeature_RNA, sample5$nFeature_RNA, sample6$nFeature_RNA,sample7$nFeature_RNA,sample8$nFeature_RNA,sample9$nFeature_RNA)

# Create dataframe
df <- data.frame(list = unlist(counts_list), Group = rep (c('1','2','3','4','5','6','7','8','9'), times = sapply(counts_list, length)))

# Plot boxplot with ggplot2
library(ggplot2)
bp_gene <- ggplot(df,aes(Group,list))+geom_boxplot() + ylab('number of genes') + xlab('sample')
bp_gene
ggsave(plot = bp_gene, width = 8,height = 6,filename = 'mouse_boxplot_gene.pdf')
```


## load the dataset
```{r}
load_seurat <- function(path,name) {
  data <- Read10X(data.dir = path)
  seurat <- CreateSeuratObject(counts = data, project = name, min.cells = 3, min.features = 200)
  return(seurat)
}
```

```{r}
sample1 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409869_78/filtered_feature_bc_matrix",'sample1')
sample2 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409870_80/filtered_feature_bc_matrix",'sample2')
sample3 <- load_seurat("~/Desktop/count_matrix_mouse/cellranger/SRR16409871_81/filtered_feature_bc_matrix",'sample3')
```

## Merge the data
```{r}
mouse.merge <- merge(sample1, y = c(sample2, sample3), add.cell.ids = c("1", "2", "3"), project = "mouse")
```

```{r}
# add the sample column
# modfiy the sample names
mouse.merge$sample <- mouse.merge$orig.ident
mouse.merge$sample <- as.factor(mouse.merge$sample)
# Assign new labels to the levels
new_levels <- c("mouse","mouse","mouse")
# Update the levels of orig.ident
levels(mouse.merge$sample) <- new_levels
```

## Quality check
```{r}
# add percentage of mitochrodial gene to metadata
mouse.merge[["percent.mt"]] <- PercentageFeatureSet(mouse.merge,pattern = '^mt-')
# visualization 
vln_mouse <- VlnPlot(mouse.merge,features = c('nCount_RNA','nFeature_RNA','percent.mt'),group.by = 'sample') 
```
Bases on the original paper and the violin plot, remove the cell with number of gene less than 1500
```{r}
# Remove the gene < 1500
# Remove mito > 20
mouse.merge[['QC']] <- ifelse(mouse.merge@meta.data$percent.mt > 20 | mouse.merge@meta.data$nFeature_RNA<1500,'Fail','Pass')
mouse.merge <- subset(mouse.merge,subset=QC=='Pass')
```

## Dimension reduction
```{r}
# Normalization and find variable genes 
mouse.merge <- NormalizeData(mouse.merge)
mouse.merge <- FindVariableFeatures(mouse.merge,selection.method = 'vst', nfeatures = 2000)
# Scale and do PCA
mouse.merge <- ScaleData(mouse.merge, verbose = FALSE)
mouse.merge <- RunPCA(mouse.merge,npcs = 30, verbose = FALSE)
```

```{r}
# Find optimal pcs
eb_mouse <- ElbowPlot(mouse.merge,ndims = 30)
eb_mouse
```

## Clustering
Try different number of pc(see plots) according to elbow plots. here use 27 pcs to do following analysis 
```{r}
# Find clusters
mouse.merge <- RunUMAP(mouse.merge,reduction = 'pca', dims = 1:27)
mouse.merge <- FindNeighbors(mouse.merge,reduction = 'pca',dims = 1:27)
mouse.merge <- FindClusters(mouse.merge,resolution = 0.8)
```

```{r}
#Visualization
cl_mouse <- DimPlot(mouse.merge,reduction = 'umap', label = TRUE, repel = TRUE,group.by = 'seurat_clusters')
```


## Cell type annotation
according to feature plots
```{r}
mouse.merge$celltype <- mouse.merge$seurat_clusters
mouse.merge$celltype <- as.factor(mouse.merge$celltype)
new_ct <- c('EX','EX','EX','EX','EX','EX','IN','EX','EX','EX','IN','EX','EX','IN','IN','EX','IN','EX','EX','EX','EX','IN','IN','EX','IN','IN','IN','IN','IN')
levels(mouse.merge$celltype) <- new_ct
```


```{r}
ct_mouse <- DimPlot(mouse.merge,reduction = 'umap', label = TRUE, repel = TRUE,group.by = 'celltype')+theme_classic(base_size = 20)
```

## Sub-clustering 
```{r}
Sub_set <- function(dataset, celltypename) {
  ct <- as_tibble(dataset@meta.data) %>% dplyr::select(seurat_clusters,celltype) %>% unique() %>% arrange(seurat_clusters) %>% filter(celltype %in% c(celltypename))
  return(ct)
}
m.sub.ex <- Sub_set(mouse.merge,'EX') 
ex.mouse <- subset(mouse.merge, idents = m.sub.ex$seurat_clusters)
m.sub.in <- Sub_set(mouse.merge,'IN') 
in.mouse <- subset(mouse.merge, idents = m.sub.in$seurat_clusters)
```